% !TEX root = ../main.tex

\chapter{Deeper analysis}
\label{chap:Deeper analysis}

First we are going to analise problem deeply without presuming any independence or probability distribution on random variables. Later in subsequent chapters we are going to focus more on where demand at time $t$ has independent Gaussian distribution and mean.

\section{Cost function}
\begin{align*}
\shortintertext{The minimizing function is:}
    & \operatorname{E} \left[
        \mathbf{s}^\intercal\mathbf{x} + \sum{c(t)}
    \right] \\
\shortintertext{Due to linerality of expectation and $x$ being variable it's equal to:}
    & \mathbf{x} \operatorname{E} \left[
        \mathbf{s}^\intercal
        \right] +
        \operatorname{E} \left[ \sum{c(t)} \right] \\
\end{align*}

Therefore only needed modeling information for supply cost is its expectation $\operatorname{E} \left[ \mathbf{s} \right]$. The other part is more trickier since $D_i$ and $D_j$ aren't usually independent.

\section{Handling demand cost non-linearity}
\label{sec:Handling demand cost non-linearity}

As we can see in equation~\ref{eq:cost-t} we have non-linearity depending whether we're satifying all demand or are we backlogging demand at time $t$. Therefore here are two possible solutions for minimizing objective function~\ref{eq:cost-f}.

\subsubsection{Simulation}
\label{subs:Simulation}

We generate multiple scenarios for demand vector, $d$ according to probability distribution. For small $n$ and relatively small number of outcomes in each random variable we can exhaustedly model each scenario, scale it appropriately and feed to MIP solver\footnote{There's a trick on using binary variable for discontinuity in cost function~\ref{eq:cost-t}}

\subsubsection{Safety net approach}
\label{subs:Safety net approach}

Alternatively, we can artificially add new constraints and avoiding backlogging with arbitrary probably. This model assumes backlogging cost are significantly greater than storage cost, that is backlogging penalty is severe.

Thus we chose values arbitrary realizations of random variables $D_t$ and add additional constraints of the form:

\begin{equation}
    \label{eq:new-cost-t}
    X_t \ge D_i \; \forall t
\end{equation}

which reduces cost function~\ref{eq:cost-t} to:
\[
c(t) = h \left( X_i - D_i \right) \; X_i \ge D_i
\]

and enables are faster solving approaches. Given parameters $D_i$ chosen and their underlying distribution we derive non-increasing function $p(t)$ which represent probability of not breaking newly introduced constraints~\ref{eq:new-cost-t}

\subsubsection{Min-cost max flow with uncertaintiy}
\label{subs:Min-cost max flow with uncertaintiy}

Since deterministic case can be modeled with min-cost max flow, we can investigate further in this directi

\subsubsection{Stochastic Dynamic Programmnig}
\label{subs:SDP}

TODO.. maybe, I've got recommendation by a friend, need to read some papers.
